Diversity dynamics in Nymphalidae butterflies: Effect of phylogenetic uncertainty on diversification rate shift estimates

### 2014-10-29 MS_revised.tex
fixed special characters

\subsection*{Abstract}

@@

\subsection*{Introduction}

@@

\subsection*{Materialas and Methods}

\subsubsection*{Data}

@@

\subsubsection*{Analyses of
Diversification}

@@

\subsubsection*{Detecting diversification shifts on phylogenetic
trees}

@@

\subsubsection*{Estimation of trait-dependent speciation
rates}

@@

\subsection*{Results}

\subsubsection*{Detecting diversification shifts on the maximum clade
credibility
tree}

@@

\subsubsection*{Phylogenetic uncertainty in the MultiMEDUSA
approach}

@@

\subsection*{Discussion}

\subsubsection*{Effects of phylogenetic uncertainty on the performance of
MEDUSA}

@@

\subsubsection*{Hostplant use and diversification in
Nymphalidae}

\paragraph*{Ithomiini}

@@

The diffuse cospeciation hypothesis predicts almost identical ages of
insects and their hostplants, while the ``resource abundance-dependent
diversity'' and the ``escape-and-radiate'' hypotheses state that insects
diversify after their hostplants {[}2--4{]}. Wheat et al. (2007)
{[}47{]} found strong evidence for a model of speciation congruent with
Ehrlich and Raven's hypothesis in Pieridae butterflies due to, in
addition to the identification of a key innovation, a burst of
diversification in glucosinolate-feeding taxa shortly afterwards (with a
lag of \textasciitilde{}10 MY). According to a recent dated phylogeny of
the Angiosperms {[}49{]}, the family Solanaceae split from its sister
group about 59 (49--68) MYA and diversification started (crown group
age) around 37 (29--47) MYA. Wahlberg et al. (2009) {[}13{]} give the
ages for origin and diversification for Ithomiini at 45 (39--53) and 37
(32--43) MYA, respectively. Thus, current evidence shows that Solanaceae
and Ithomiini might have diversified around the same time, during the
Late Eocene and Oligocene, and this would be congruent with the diffuse
cospeciation hypothesis.

\paragraph*{Danaini}
Our MultiMEDUSA approach showed a significant slowdown in net
diversification rate in the subtribe Danaina of the Danini. Both Danaina
and the sister clade Euploeina feed mainly on Apocynaceae and thus a
hostplant shift should not be responsible for the observed slowdown of
diversification in the Danaina. As expected, our BiSSE analysis of
Apocynaceae feeders shows that there is no effect of feeding on this
plant family on the net diversification rates of Nymphalidae lineages.
Many of the Danaina are large, strong fliers, highly migratory and
involved in mimicry rings. Among them is for example the monarch
(\emph{Danaus plexippus}), probably the most well known of all migratory
butterflies. The causes for a lower net diversification rate in the
Danaina remains to be investigated, but their great dispersal power
might be involved in preventing allopatric speciation. It has been found
in highly vagile species in the nymphalid genus \emph{Vanessa} that
dispersal has homogenized populations due to gene flow, as old and
vagile species seem to be genetically homogeneous while younger
widespread species show higher genetic differentiation in their
populations {[}50{]}.

\paragraph*{Satyrini}

@@

\subsection*{Conclusions}
We found that even though MEDUSA estimated several diversification
shifts in the maximum clade credibility tree of Nymphalidae, only a few
of these splits were found in more than 90\% of the trees from the
posterior distribution. In the literature, it is common practice that
conclusions are based on the splits estimated on the maximum clade
credibility tree. However, by using a MultiMEDUSA approach, we found
that for this Nymphalidae dataset some of these splits might be greatly
affected by phylogenetic uncertainty. Moreover, some of these splits can
be recovered either as increases or decreases in net diversification
rate depending on the tree from the posterior distribution that was used
for analysis. This means that contradictory conclusions would be made if
only the maximum clade credibility tree was used for analysis. We
recommend that all datasets should be analyzed using both approaches,
MEDUSA and MultiMEDUSA, in order to test whether the results are robust
when phylogenetic uncertainty is taken into account.

@@

\subsection*{Acknowledgments}
We are thankful to Mark Cornwall for help with the script to extend
MEDUSA to include phylogenetic uncertainty, Niklas Wahlberg for
commenting on the manuscript and giving us the posterior distribution of
trees, Luke Harmon for commenting on the manuscript and anonymous
reviewers for their comments, which greatly improved the manuscript,
Jessica Slove Davidson and Niklas Janz for access to their hostplant
data. The study was supported by a Kone Foundation grant (awarded to
Niklas Wahlberg), Finland (C. Pe\~na) and the Research Council of Norway
(grant no. 204308 to M. Espeland). We acknowledge CSC--IT Center for
Science Ltd. (Finland) for the allocation of computational resources.

@@

\section*{Supporting Information Legends}
\begin{description}
 \item {\bf Supporting\_Information\_S01.nex.} MCC Nymphalidae tree from
{[}13{]}

 \item {\bf Supporting\_Information\_S02.R} R script to run MEDUSA on the
MCC Nymphalidae tree from {[}13{]}. This script removes the outgroup
taxa and loads the richness data for the tree terminals.

 \item {\bf Supporting\_Information\_S03.csv} Species richness for lineages
in Nymphalidae.

 \item {\bf Supporting\_Information\_S04.txt} Results of running MEDUSA on
the MCC tree.

 \item {\bf Supporting\_Information\_S05\_1000\_random\_trees\_no\_outgroups.zip} 1000 random trees from {[}13{]}.

 \item {\bf Supporting\_Information\_S06.R} R script to run a MultiMEDUSA
analysis on 1000 random trees from {[}13{]}.

 \item {\bf Supporting\_Information\_S07\_multimedusa\_on\_1000\_trees.zip} Raw results from the MultiMEDUSA run on the random sample of trees from
the posterior distribution.

 \item {\bf Supporting\_Information\_S08.pdf} Summary of the MulitMEDUSA
analysis on the 1000 trees from the posterior distribution.

 \item {\bf Supporting\_Information\_S09.tree} MCC tree from the 1000
random trees selected from the posterior distribution.

 \item {\bf Supporting\_Information\_S10.R} R script to run a MEDUSA
analysis on the MCC tree from the 1000 random trees selected from the
posterior distribution.

 \item {\bf Supporting\_Information\_S11.txt} Summary results from a MEDUSA
run on the MCC tree from the 1000 random trees selected from the
posterior distribution.

 \item {\bf Supporting\_Information\_S12.pdf} Figure for MEDUSA run on MCC
tree from random 1000 trees.

 \item {\bf Supporting\_Information\_S13.txt} Summary results from a
MultiMEDUSA run on the 1000 random trees selected from the posterior
distribution.

 \item {\bf Supporting\_Information\_S14.pdf} Probability of nodes with
estimated rates from a MultiMEDUSA run on the 1000 random trees selected
from the posterior distribution.

 \item {\bf Supporting\_Information\_S15.csv} Hostplants of Nymphalidae
butterflies recorded from the literature.

 \item {\bf Supporting\_Information\_S16.csv} References for hostplants
data.

 \item {\bf Supporting\_Information\_S17.csv} Data matrix with character
states for hosplant use.

 \item {\bf Supporting\_Information\_S18.R} R script for running the BiSSE
analysis.

 \item {\bf Supporting\_Information\_S19.csv} Raw results for the BiSSE
analysis.

 \item {\bf Supporting\_Information\_S20.csv} Raw results for the combined
BiSSE analysis.

@@

2. Nyman T, Linder HP, Peña C, Malm T, Wahlberg N (2012) Climate-driven
diversity dynamics in plants and plant-feeding insects. Ecol Lett 15:
889--898.
doi:\href{http://dx.doi.org/10.1111/j.1461-0248.2012.01782.x}{10.1111/j.1461-0248.2012.01782.x}.

@@

10. Heikkilä M, Kaila L, Mutanen M, Peña C, Wahlberg N (2012) Cretaceous
origin and repeated tertiary diversification of the redefined
butterflies. P Roy Soc B 279: 1093--1099.
doi:\href{http://dx.doi.org/10.1098/rspb.2011.1430}{10.1098/rspb.2011.1430}.

11. Elias M, Joron M, Willmott KR, Silva-Brandão KL, Kaiser V, et al.
(2009) Out of the Andes: Patterns of diversification in clearwing
butterflies. Mol Ecol 18: 1716--1729.
doi:\href{http://dx.doi.org/10.1111/j.1365-294X.2009.04149.x}{10.1111/j.1365-294X.2009.04149.x}.

@@

13. Wahlberg N, Leneveu J, Kodandaramaiah U, Peña C, Nylin S, et al.
(2009) Nymphalid butterflies diversify following near demise at the
Cretaceous/Tertiary boundary. P Roy Soc B 276: 4295--4302.
doi:\href{http://dx.doi.org/10.1098/rspb.2009.1303}{10.1098/rspb.2009.1303}.

14. Pe\~na C, Wahlberg N (2008) Prehistorical climate change increased
diversification of a group of butterflies. Biol Lett 4: 274--278.
doi:\href{http://dx.doi.org/10.1098/rsbl.2008.0062}{10.1098/rsbl.2008.0062}.

15. Ferrer-Paris JR, S{á}nchez-Mercado A, Viloria ÁL, Donaldson J (2013)
Congruence and Diversity of Butterfly-Host Plant Associations at Higher
Taxonomic Levels. PloS one 8: e63570.
doi:\href{http://dx.doi.org/0.1371/journal.pone.0063570}{0.1371/journal.pone.0063570}.

@@

20. Matos-Maraví PF, Peña C, Willmott KR, Freitas AV, Wahlberg N (2013)
Systematics and evolutionary history of butterflies in the ``Taygetis
clade'' (Nymphalidae: Satyrinae: Euptychiina): Towards a better
understanding of Neotropical biogeography. Mol Phylogenet Evol 66:
54--68.
doi:\href{http://dx.doi.org/10.1016/j.ympev.2012.09.005}{10.1016/j.ympev.2012.09.005}.

21. Brower AVZ, Wahlberg N, Ogawa JR, Boppré M, Vane-Wright RI (2010)
Phylogenetic relationships among genera of danaine butterflies
(Lepidoptera: Nymphalidae) as implied by morphology and DNA sequences.
Syst Biodivers 8: 75--89.
doi:\href{http://dx.doi.org/10.1080/14772001003626814}{10.1080/14772001003626814}.

22. Kodandaramaiah U, Lees DC, Müller CJ, Torres E, Karanth KP, et al.
(2010) Phylogenetics and biogeography of a spectacular Old World
radiation of butterflies: the subtribe Mycalesina (Lepidoptera:
Nymphalidae: Satyrini). BMC Evol Biol 10: 172.
doi:\href{http://dx.doi.org/10.1186/1471-2148-10-172}{10.1186/1471-2148-10-172}.

23. Kodandaramaiah U, Peña C, Braby MF, Grund R, Müller CJ, et al.
(2010) Phylogenetics of Coenonymphina (Nymphalidae: Satyrinae) and the
problem of rooting rapid radiations. Mol Phylogenet Evol 54: 386--394.
doi:\href{http://dx.doi.org/10.1016/j.ympev.2009.08.012}{10.1016/j.ympev.2009.08.012}.

@@

27. Peña C, Wahlberg N, Weingartner E, Kodandaramaiah U, Nylin S, et al.
(2006) Higher level phylogeny of Satyrinae butterflies (Lepidoptera:
Nymphalidae) based on DNA sequence data. Mol Phylogenet Evol 40: 29--49.
doi:\href{http://dx.doi.org/10.1016/j.ympev.2006.02.007}{10.1016/j.ympev.2006.02.007}.

@@

29. Silva-Brandão KL, Wahlberg N, Francini RB, Azeredo-Espin AML, Brown
Jr. KS, et al. (2008) Phylogenetic relationships of butterflies of the
tribe Acraeini (Lepidoptera, Nymphalidae, Heliconiinae) and the
evolution of host plant use. Mol Phylogenet Evol 46: 515--531.
doi:\href{http://dx.doi.org/10.1016/j.ympev.2007.11.024}{10.1016/j.ympev.2007.11.024}.

30. Peña C, Nylin S, Wahlberg N (2011) The radiation of Satyrini
butterflies (Nymphalidae: Satyrinae): a challenge for phylogenetic
methods. Zool J Linn Soc 161: 64--87.
doi:\href{http://dx.doi.org/10.1111/j.1096-3642.2009.00627.x}{10.1111/j.1096-3642.2009.00627.x}.

31. Peña C, Nylin S, Freitas AVL, Wahlberg N (2010) Biogeographic
history of the butterfly subtribe Euptychiina (Lepidoptera, Nymphalidae,
Satyrinae). Zool Scr 39: 243--258.
doi:\href{http://dx.doi.org/10.1111/j.1463-6409.2010.00421.x}{10.1111/j.1463-6409.2010.00421.x}.

@@

34. Beccaloni G, Viloria Á, Hall S, Robinson G (2008) Catalogue of the
hostplants of the Neotropical butterflies. London: The Natural History
Museum.

@@

45. Brower AVZ, Freitas AVL, Lee M-M, Silva-Brandão KL, Whinnett A, et
al. (2006) Phylogenetic relationships among the Ithomiini (Lepidoptera:
Nymphalidae) inferred from one mitochondrial and two nuclear gene
regions. Syst Entomol 31: 288--301.
doi:\href{http://dx.doi.org/10.1111/j.1365-3113.2006.00321.x}{10.1111/j.1365-3113.2006.00321.x}.

@@

50. Wahlberg N, Rubinoff D (2011) Vagility across Vanessa (Lepidoptera:
Nymphalidae): mobility in butterfly species does not inhibit the
formation and persistence of isolated sister taxa. Syst Entomol 36:
362--370.
doi:\href{http://dx.doi.org/10.1111/j.1365-3113.2010.00566.x}{10.1111/j.1365-3113.2010.00566.x}.

>>>>

\section*{Abstract}

@@

\section*{Introduction}

@@

\section*{Materials and Methods}

\subsection*{Data}

@@

\subsection*{Analyses of
Diversification}

@@

\subsection*{Detecting diversification shifts on phylogenetic
trees}

@@

\subsection*{Estimation of trait-dependent speciation
rates}

@@

\section*{Results}

\subsection*{Detecting diversification shifts on the maximum clade
credibility tree}

@@

\subsection*{Phylogenetic uncertainty in the MultiMEDUSA
approach}

@@

\section*{Discussion}

\subsection*{Effects of phylogenetic uncertainty on the performance of
MEDUSA}

@@

\subsection*{Hostplant use and diversification in
Nymphalidae}

\subsubsection*{Ithomiini}

@@

The diffuse cospeciation hypothesis predicts almost identical ages of
insects and their hostplants, while the ``resource abundance-dependent
diversity'' and the ``escape-and-radiate'' hypotheses state that insects
diversify after their hostplants {[}2--4{]}. Wheat et al. (2007)
{[}47{]} found strong evidence for a model of speciation congruent with
Ehrlich and Raven's hypothesis in Pieridae butterflies due to, in
addition to the identification of a key innovation, a burst of
diversification in glucosinolate-feeding taxa shortly afterwards (with a
lag of circa 10 MY). According to a recent dated phylogeny of
the Angiosperms {[}49{]}, the family Solanaceae split from its sister
group about 59 (49--68) MYA and diversification started (crown group
age) around 37 (29--47) MYA. Wahlberg et al. (2009) {[}13{]} give the
ages for origin and diversification for Ithomiini at 45 (39--53) and 37
(32--43) MYA, respectively. Thus, current evidence shows that Solanaceae
and Ithomiini might have diversified around the same time, during the
Late Eocene and Oligocene, and this would be congruent with the diffuse
cospeciation hypothesis.

\subsubsection*{Danaini}
Our MultiMEDUSA approach showed a significant slowdown in net
diversification rate in the subtribe Danaina of the Danini. Both Danaina
and the sister clade Euploeina feed mainly on Apocynaceae and thus a
hostplant shift should not be responsible for the observed slowdown of
diversification in the Danaina. As expected, our BiSSE analysis of
Apocynaceae feeders shows that there is no effect of feeding on this
plant family on the net diversification rates of Nymphalidae lineages.
Many of the Danaina are large, strong fliers, highly migratory and
involved in mimicry rings. Among them is for example the monarch
(\emph{Danaus plexippus}), probably the most well known of all migratory
butterflies. The causes for a lower net diversification rate in the
Danaina remains to be investigated, but their great dispersal power
might be involved in preventing allopatric speciation. It has been found
in highly vagile species in the nymphalid genus \emph{Vanessa} that
dispersal has homogenized populations due to gene flow, as old and
vagile species seem to be genetically homogeneous while younger
widespread species show higher genetic differentiation in their
populations {[}50{]}.

\subsubsection*{Satyrini}

@@

\section*{Conclusions}
We found that even though MEDUSA estimated several diversification
shifts in the maximum clade credibility tree of Nymphalidae, only a few
of these splits were found in more than 90\% of the trees from the
posterior distribution. In the literature, it is common practice that
conclusions are based on the splits estimated on the maximum clade
credibility tree. However, by using a MultiMEDUSA approach, we found
that for this Nymphalidae dataset some of these splits might be greatly
affected by phylogenetic uncertainty. Moreover, some of these splits can
be recovered either as increases or decreases in net diversification
rate depending on the tree from the posterior distribution that was used
for analysis. This means that contradictory conclusions would be made if
only the maximum clade credibility tree was used for analysis. We
recommend that all datasets should be analyzed using both approaches,
MEDUSA and MultiMEDUSA, in order to test whether the results are robust
when phylogenetic uncertainty is taken into account.

@@

\section*{Acknowledgments}
We are thankful to Mark Cornwall for help with the script to extend
MEDUSA to include phylogenetic uncertainty, Niklas Wahlberg for
commenting on the manuscript and giving us the posterior distribution of
trees, Luke Harmon for commenting on the manuscript and anonymous
reviewers for their comments, which greatly improved the manuscript,
Jessica Slove Davidson and Niklas Janz for access to their hostplant
data. The study was supported by a Kone Foundation grant (awarded to
Niklas Wahlberg), Finland (C. Pe\~na) and the Research Council of Norway
(grant no. 204308 to M. Espeland). We acknowledge CSC--IT Center for
Science Ltd. (Finland) for the allocation of computational resources.

@@

\section*{Supporting Information Legends}
\begin{description}
 \item {\bf Supporting Information S01.nex.} MCC Nymphalidae tree from
{[}13{]}

 \item {\bf Supporting Information S02.R} R script to run MEDUSA on the
MCC Nymphalidae tree from {[}13{]}. This script removes the outgroup
taxa and loads the richness data for the tree terminals.

 \item {\bf Supporting Information S03.csv} Species richness for lineages
in Nymphalidae.

 \item {\bf Supporting Information S04.txt} Results of running MEDUSA on
the MCC tree.

 \item {\bf Supporting Information S05 1000 random trees no outgroups.zip} 1000 random trees from {[}13{]}.

 \item {\bf Supporting Information S06.R} R script to run a MultiMEDUSA
analysis on 1000 random trees from {[}13{]}.

 \item {\bf Supporting Information S07 multimedusa on 1000 trees.zip} Raw results from the MultiMEDUSA run on the random sample of trees from
the posterior distribution.

 \item {\bf Supporting Information S08.pdf} Summary of the MulitMEDUSA
analysis on the 1000 trees from the posterior distribution.

 \item {\bf Supporting Information S09.tree} MCC tree from the 1000
random trees selected from the posterior distribution.

 \item {\bf Supporting Information S10.R} R script to run a MEDUSA
analysis on the MCC tree from the 1000 random trees selected from the
posterior distribution.

 \item {\bf Supporting Information S11.txt} Summary results from a MEDUSA
run on the MCC tree from the 1000 random trees selected from the
posterior distribution.

 \item {\bf Supporting Information S12.pdf} Figure for MEDUSA run on MCC
tree from random 1000 trees.

 \item {\bf Supporting Information S13.txt} Summary results from a
MultiMEDUSA run on the 1000 random trees selected from the posterior
distribution.

 \item {\bf Supporting Information S14.pdf} Probability of nodes with
estimated rates from a MultiMEDUSA run on the 1000 random trees selected
from the posterior distribution.

 \item {\bf Supporting Information S15.csv} Hostplants of Nymphalidae
butterflies recorded from the literature.

 \item {\bf Supporting Information S16.csv} References for hostplants
data.

 \item {\bf Supporting Information S17.csv} Data matrix with character
states for hosplant use.

 \item {\bf Supporting Information S18.R} R script for running the BiSSE
analysis.

 \item {\bf Supporting Information S19.csv} Raw results for the BiSSE
analysis.

 \item {\bf Supporting Information S20.csv} Raw results for the combined
BiSSE analysis.

@@

2. Nyman T, Linder HP, Pe{\~n}a C, Malm T, Wahlberg N (2012) Climate-driven
diversity dynamics in plants and plant-feeding insects. Ecol Lett 15:
889--898.
doi:\href{http://dx.doi.org/10.1111/j.1461-0248.2012.01782.x}{10.1111/j.1461-0248.2012.01782.x}.

@@

10. Heikkil{\"a} M, Kaila L, Mutanen M, Pe{\~n}a C, Wahlberg N (2012) Cretaceous
origin and repeated tertiary diversification of the redefined
butterflies. P Roy Soc B 279: 1093--1099.
doi:\href{http://dx.doi.org/10.1098/rspb.2011.1430}{10.1098/rspb.2011.1430}.

11. Elias M, Joron M, Willmott KR, Silva-Brand{\~a}o KL, Kaiser V, et al.
(2009) Out of the Andes: Patterns of diversification in clearwing
butterflies. Mol Ecol 18: 1716--1729.
doi:\href{http://dx.doi.org/10.1111/j.1365-294X.2009.04149.x}{10.1111/j.1365-294X.2009.04149.x}.

@@

13. Wahlberg N, Leneveu J, Kodandaramaiah U, Pe{\~n}a C, Nylin S, et al.
(2009) Nymphalid butterflies diversify following near demise at the
Cretaceous/Tertiary boundary. P Roy Soc B 276: 4295--4302.
doi:\href{http://dx.doi.org/10.1098/rspb.2009.1303}{10.1098/rspb.2009.1303}.

14. Pe{\~n}a C, Wahlberg N (2008) Prehistorical climate change increased
diversification of a group of butterflies. Biol Lett 4: 274--278.
doi:\href{http://dx.doi.org/10.1098/rsbl.2008.0062}{10.1098/rsbl.2008.0062}.

15. Ferrer-Paris JR, S{\'a}nchez-Mercado A, Viloria {\'A}L, Donaldson J (2013)
Congruence and Diversity of Butterfly-Host Plant Associations at Higher
Taxonomic Levels. PloS one 8: e63570.
doi:\href{http://dx.doi.org/0.1371/journal.pone.0063570}{0.1371/journal.pone.0063570}.

@@

20. Matos-Marav{\'i} PF, Pe{\~n}a C, Willmott KR, Freitas AV, Wahlberg N (2013)
Systematics and evolutionary history of butterflies in the ``Taygetis
clade'' (Nymphalidae: Satyrinae: Euptychiina): Towards a better
understanding of Neotropical biogeography. Mol Phylogenet Evol 66:
54--68.
doi:\href{http://dx.doi.org/10.1016/j.ympev.2012.09.005}{10.1016/j.ympev.2012.09.005}.

21. Brower AVZ, Wahlberg N, Ogawa JR, Boppr{\'e} M, Vane-Wright RI (2010)
Phylogenetic relationships among genera of danaine butterflies
(Lepidoptera: Nymphalidae) as implied by morphology and DNA sequences.
Syst Biodivers 8: 75--89.
doi:\href{http://dx.doi.org/10.1080/14772001003626814}{10.1080/14772001003626814}.

22. Kodandaramaiah U, Lees DC, M{\"u}ller CJ, Torres E, Karanth KP, et al.
(2010) Phylogenetics and biogeography of a spectacular Old World
radiation of butterflies: the subtribe Mycalesina (Lepidoptera:
Nymphalidae: Satyrini). BMC Evol Biol 10: 172.
doi:\href{http://dx.doi.org/10.1186/1471-2148-10-172}{10.1186/1471-2148-10-172}.

23. Kodandaramaiah U, Pe{\~n}a C, Braby MF, Grund R, M{\"u}ller CJ, et al.
(2010) Phylogenetics of Coenonymphina (Nymphalidae: Satyrinae) and the
problem of rooting rapid radiations. Mol Phylogenet Evol 54: 386--394.
doi:\href{http://dx.doi.org/10.1016/j.ympev.2009.08.012}{10.1016/j.ympev.2009.08.012}.

@@

27. Pe{\~n}a C, Wahlberg N, Weingartner E, Kodandaramaiah U, Nylin S, et al.
(2006) Higher level phylogeny of Satyrinae butterflies (Lepidoptera:
Nymphalidae) based on DNA sequence data. Mol Phylogenet Evol 40: 29--49.
doi:\href{http://dx.doi.org/10.1016/j.ympev.2006.02.007}{10.1016/j.ympev.2006.02.007}.

@@

29. Silva-Brand{\~a}o KL, Wahlberg N, Francini RB, Azeredo-Espin AML, Brown
Jr. KS, et al. (2008) Phylogenetic relationships of butterflies of the
tribe Acraeini (Lepidoptera, Nymphalidae, Heliconiinae) and the
evolution of host plant use. Mol Phylogenet Evol 46: 515--531.
doi:\href{http://dx.doi.org/10.1016/j.ympev.2007.11.024}{10.1016/j.ympev.2007.11.024}.

30. Pe{\~n}a C, Nylin S, Wahlberg N (2011) The radiation of Satyrini
butterflies (Nymphalidae: Satyrinae): a challenge for phylogenetic
methods. Zool J Linn Soc 161: 64--87.
doi:\href{http://dx.doi.org/10.1111/j.1096-3642.2009.00627.x}{10.1111/j.1096-3642.2009.00627.x}.

31. Pe{\~n}a C, Nylin S, Freitas AVL, Wahlberg N (2010) Biogeographic
history of the butterfly subtribe Euptychiina (Lepidoptera, Nymphalidae,
Satyrinae). Zool Scr 39: 243--258.
doi:\href{http://dx.doi.org/10.1111/j.1463-6409.2010.00421.x}{10.1111/j.1463-6409.2010.00421.x}.

@@

34. Beccaloni G, Viloria {\'A}, Hall S, Robinson G (2008) Catalogue of the
hostplants of the Neotropical butterflies. London: The Natural History
Museum.

@@

45. Brower AVZ, Freitas AVL, Lee M-M, Silva-Brand{\~a}o KL, Whinnett A, et
al. (2006) Phylogenetic relationships among the Ithomiini (Lepidoptera:
Nymphalidae) inferred from one mitochondrial and two nuclear gene
regions. Syst Entomol 31: 288--301.
doi:\href{http://dx.doi.org/10.1111/j.1365-3113.2006.00321.x}{10.1111/j.1365-3113.2006.00321.x}.

@@

50. Wahlberg N, Rubinoff D (2011) Vagility across \emph{Vanessa} (Lepidoptera:
Nymphalidae): mobility in butterfly species does not inhibit the
formation and persistence of isolated sister taxa. Syst Entomol 36:
362--370.
doi:\href{http://dx.doi.org/10.1111/j.1365-3113.2010.00566.x}{10.1111/j.1365-3113.2010.00566.x}.
